---
title: "Bayesian modeling and prediction for movies by Kunyu HE"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
    pandoc_args: [
      "--number-sections",
    ]
---

As an assumed data scientist at *Paramount Pictures*, I need to work on the data set `movies` to learn what attributes make a movie popular.

__Setup__

__Load packages__

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(BAS)
library(MASS)
library(GGally)
```

__Load data__

```{r load-data}
load("movies.Rdata")
```



* * *

# Part 1: Data


* * *

```{r data_structure}
str(movies)
```

The data set `movies` is comprised of 651 observations of 32 variables. The observations are randomly sampled movies produced and released before 2016 from *Rotten Tomatoes* and *IMDB* [1].

*([1]Source: Project Files and Rubric, https://www.coursera.org/learn/linear-regression-model/peer/gIqqL/data-analysis-project, accessed at 09/27/2017)*

__As the observations were collected with random sampling, we can generalize our inference to a population of all the movies on *Rotten Tomatoes* and *IMDB*, which were produced and released before 2016.__ However, these websites collect ratings from voluntary behaviours of viewers, therefore the ratings could be biased initially.

__As the collection is observational, we can only conclude on associations among variables.__ 

* * *

# Part 2: Data Manipulation

* * *

Calculate following variables for future use before EDA process:

`feature_film`: indicates whether or it is a feature film;
`drama`: indicates whether the film genre is a drama;
`mpaa_rating_R`: indiactes whether the film was rated "R" or not;
`oscar_season`: movie was realeased in October, November, or December.
`summer_season`: movie was released in May, June, July, or August.

```{r calculate_variables}
movies$feature_film <- ifelse(movies$title_type == 'Feature Film', 'yes', 'no')
movies$drama <- ifelse(movies$genre == 'Drama', 'yes', 'no')
movies$mpaa_rating_R <- ifelse(movies$mpaa_rating == 'R', 'yes', 'no')
movies$oscar_season <- sapply( movies$thtr_rel_month, function(value) if ( value >= 10 ) "yes" else "no" )
movies$summer_season <- sapply( movies$thtr_rel_month, function(value) if ( value >= 5 & value <= 8 ) "yes" else "no" )
```


* * *

# Part 3: Exploratory Data Analysis

* * *

Before EDA, narrow down the data set to a sub-set with the variables of interest and omit the `"NA"` value.

```{r sub_omit}
movies_BA <- na.omit(movies[c("audience_score", "feature_film", "drama", "mpaa_rating_R", "runtime", "thtr_rel_year", "oscar_season", "summer_season", "best_pic_nom", "best_pic_win", "best_actor_win", "best_dir_win", "best_actress_win", "imdb_rating", "imdb_num_votes", "critics_score", "top200_box")])
```


* * *

## Response Variable: `audience_score`

* * *

In order to develop a Bayesian regression model to predict `audience_score`, which implies audience ratings on *Rotten Tomatoes*, first we need to scrutinize the dependent variable. Print a summary table and a frequency distribution histogram for it.

```{r dependent_variable}
summary(movies_BA$audience_score)
ggplot(data=movies_BA, aes(x=audience_score))+geom_histogram(binwidth = 3, fill="lightcoral", color="Brown")+xlab("Audience Score") + ggtitle("Frequency Distribution of Rotten Tomatoes Audience Score") + theme(plot.title = element_text(hjust=0.5, vjust=3))
```

THe sample distribution is bimodal and left-skewed, with a median larger than the sample mean by approximately 3 points. However, as sample size is much larger than 30, we can assume its distribution to be nearly normal.

* * *

## `audience_score` against `feature_film`

* * *

Scrutinize the association between `audience_score` and `feature_film`. As the predictor is categorical and response variable is numerical, use a box plot.

```{r score_feature}
movies_BA$feature_film <- as.factor(movies_BA$feature_film)
movies_BA$feature_film <- relevel(movies_BA$feature_film, ref="yes")
ggplot(movies_BA, aes(x=feature_film, y=audience_score, fill = feature_film)) + geom_boxplot(outlier.color = "purple") + ggtitle("Audience Score versus Feature Film") + labs(x="Feature Film", y="Audience Score") + theme(plot.title = element_text(hjust=0.5, vjust=3))
movies_BA %>% 
  group_by(feature_film) %>%
  summarise(count=n())
```

__Movies that are not feature films appear to have higher audience scores.__ IQR of the non-feature films lies almost entirely above that of feature films on the scale of popularity. Variation in audience score of the non-feature films is much lower, and there are only few ouliers in the group rated extremely low. As the summary tables above shows, there are 59 observations in the non-feature group, which is significantly large to support our impression. 

* * *

## `audience_score` against `imdb_rating`

* * *

Collected from another website *IMDB*, `imdb_rating` is a numerical variable and another proxy for the popularity of particular movies. Print a scatter plot and the correlation coefficient to find potential correlation between both variables.

```{r rotten_imdb}
ggplot(movies_BA, aes(y=audience_score, x=imdb_rating))+geom_point(size=1, shape=3, color="mediumorchid4")+ylab("Audience Score")+xlab("IMDB Rating") + ggtitle(" Audience Score against IMDB Rating") + theme(plot.title=element_text(hjust=0.5, vjust=3))
cor(movies_BA$audience_score, movies_BA$imdb_rating)
```

__Seems audience voting on *IMDB* and *Rotten Tomatoes* somehow share views about particular films.__ From the scatter plot above, we expect a potential positive correlation between `audience_score` and `imdb_rating`. The correlation coefficient is 0.86, which is large enough to support a strong correlation. 

* * *

# Part 4: Modeling

* * *

* * *

## Model with BIC prior

* * *

Construct a Bayesian model with BIC model selection method, where lowest BIC indicates the best fit.

```{r BIC_model}
BIC = bas.lm(audience_score ~ ., data = movies_BA, prior = "BIC", modelprior = uniform())
summary(BIC)
```

Plot model diagnostics and model rank image for the model with BIC prior.

```{r resid_BIC}
plot(BIC)
image(BIC, rotate=FALSE, cex.lab= 0.01)
```

There are 3 Outliers in the plot: 123, 216 and 231. The distribution of `"Residuals vs Fitted"` does not appear to be random, which implies the model shouldn't be re-addressed. 

In `"Model probabilies"`, probabilities tend to increase along with the search order, which implies a trend of time.

Furthermore, `"Model complexity"` is separated into three distinct sections, possibly resulting from significant effects that some categorical variables in the model take.

__Finally, the `"Inclusion Probabilties"` and model rank suggest a different model, where three variables are listed as potential predictors of `audience_score`: `runtime`, `imdb_rating`, and `critics_score`; with `imdb_rating` and `critics_score` as consistent predictors cross all models.__

* * *

## Model with Zellner's g prior

* * *

For comparison, construct and select another model with Zellner's g prior, full.

```{r Zs_full_model}
Zs_full = bas.lm(audience_score ~ ., data = movies_BA, prior = "ZS-full", modelprior = uniform())
summary(Zs_full)
```

Plot model diagnostics and model rank image for the model with Zellner's g prior, full.

```{r resid_Zs_full}
plot(Zs_full)
image(Zs_full, rotate=FALSE, cex.lab= 0.01)
```


Likewise, there are 3 Outliers in the plot: 123, 216 and 231. The distribution of `"Residuals vs Fitted"` does not appear to be random, which implies the model shouldn't be re-addressed.

However, the `"Model Probabilities"` plot shows consistent cumulative model probability for each time a new model is sampled. The plot shows almost straight upward line indicating added probability as more models are added.

Furthermore, `"Model complexity"` is separated into three distinct sections, possibly resulting from significant effects that some categorical variables in the model take.

__Finally, the `"Inclusion Probabilties"` and model rank suggest a different model, where three variables are listed as potential predictors of `audience_score`: `runtime`, `imdb_rating`, and `critics_score`; with `imdb_rating` and `critics_score` as consistent predictors cross all models.__

* * *

## Final Model

* * *

Our final model would include merely three predictors, `runtime`, `imdb_rating`, and `critics_score`. Construct the model and print a summary table for both the model and its coefficients.

```{r final_model}
final_model = bas.lm(audience_score ~ runtime + imdb_rating + critics_score, data = movies_BA, prior = "BIC", modelprior = uniform())
summary(final_model)
confint(coefficients(final_model))
```

__The posterior probability of the final model is 0.453, which is significantly high.__ The formula is:

$$audience_score = 62.35 - 0.028 \cdot runtime + 14.963 \cdot imdb_rating + 0.065 \cdot critics_score$$ 

__In conclusion, ceteris paribus, we expect a 0.028 decrease of `audience_score` with a one-minute increase in `runtime`, on average; ceteris paribus, we expect a 14.963 increase of `audience_score` with a one point increase in `imdb_rating`, on average; ceteris paribus, we expect a 0.065 increase of `audience_score` with a one point increase in `critics_score`, on average.__

* * *

# Part 5: Prediction

* * *

For prediction, pick *American Assasin*[3]. Caculate by coding:

*([3]Source: American Assasin, IMDB at http://www.imdb.com/title/tt1961175/?ref_=inth_ov_tt, Rotten Tomatoes at https://www.rottentomatoes.com/m/american_assassin_2017, both accessed on 27/09/2017)*

```{r predict_AA}
AA <-data.frame(119, 6.6, 33)
names(AA)<-c("runtime", "imdb_rating", "critics_score")
predict_AA_1 <- predict(final_model,AA,estimator="BPM", se.fit=TRUE)
predict_AA_2 <- predict(final_model,AA,estimator="MPM", se.fit=TRUE)
predict_AA_1$fit
predict_AA_2$fit
```

__According to the prediction, the audience score of *American Assasin* should be 61-62, where the true score is 69. Our prediction is relatively close.__

* * *

# Part 6: Conclusion

* * *

__We conclude that based on the given data set `movies`, `critics_score`, `imdb_rating` are predictors positively correlate with `audience_score`; `runtime`, contrarily, negatively correlates with `audience_score`; and `imdb_rating` take the strongest effect on audience score of a movies on *Rotten Tomatoes* and *IMDB*, which were produced and released before 2016.__

One shortcoming of this analysis is the lack of prior model. It will greatly enhance our analysis if we research more and use a better prior model for our regression. 
